Adaboost for face image classification

### 2019-11-12 simple_ml/example/report3/paper.tex
add report3.

\makeatletter
\newif\if@restonecol
\makeatother
\let\algorithm\relax
\let\endalgorithm\relax
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}%[ruled,vlined]{
\usepackage{algpseudocode}
\usepackage{amsmath}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}  % Use Input in the format of Algorithm
\renewcommand{\algorithmicensure}{\textbf{Output:}} % Use Output in the format of Algorithm 
% Other popular packages for formatting tables and equations include:

@@

% Define document title and author
	\title{Logistic Regression and SVM}
	\maketitle

% Write abstract here
\begin{abstract}
In this report, we solve the binary classification problem in a9a dataset by Logistic Regression (LR) and Support Vector Machine (SVM). \\
We perform experiments on four aspects: \\
1. Hyper-parameter tuning in SVM. \\
2. Loss function comparision in SVM. \\
3. initializer comparision in both LR and SVM. \\
4. Optimizer comparision in both LR and SVM. \\
\end{abstract}

% Each section begins with a \section{title} command
\section{Introduction}
	% \PARstart{}{} creates a tall first letter for this first paragraph
\PARstart{L}{ogistic} Regression(LR) is an extension of linear regression, which use $\sigma$ function as activation to map the output of linear regression into a interval (0, 1). So that LR can output the probability in binary classification. SVM is also an extension of linear regression. Based on the idea "large margin classification", SVM can perform very well in many classification task. Even now, LR and SVM are widely used in academic and industry community. These two algorithms are classic, so we fully explore them by solving a9a binary classification task. Beyond that, we will explore the initializer and optimizer selection in both LR and SVM.

% Main Part
\section{Methods and Theory}
In this part, we first define logistic regression and LR's loss function Binary Cross Entrophy (BCE). Then we define SVM and SVM's loss function Hinge Loss (HL).
\subsection{Logistic Regression}
Given dataset $ D=\{x_{i}, y_{i}\}_{i=1}^{m}$, where $x_{i} \in \mathbf{R}^{n}$ and $y_{i} \in \{0, 1\}$. Mathematically, for a single training datapoint $(x, y)$, LR assumes: \par
 \begin{center}
	\begin{equation}
	P(Y = 1|X = x) = \sigma(z) 
	\end{equation}
	\begin{equation}
	P(Y = 0|X = x) = 1 - \sigma(z)
	\end{equation}
	\begin{equation}
	z = F(x; W)
	\end{equation}
\end{center} \par
where $F(x; W)$ is linear regression parameterized by $W$ and $\sigma(.)$ is sigmoid function, where $\sigma(z) = \frac{1}{1 + e^{-z}} \in (0, 1)$.
\subsection{Binary Cross Entrophy}
Logistic regression assumes that each datapoint is independent, so the likelihood of all data is: \par
\begin{equation}
	\begin{aligned}
	L(W) &= \prod_{i=1}^m P(Y = y_i | X = x_i) \\ &= \prod_{i=1}^m \sigma(z)^{y_i} (1 - \sigma(z)^{1 - y_i} )
	\end{aligned}
\end{equation} \par

So our goal is to maximize the likelihood of data, in practice we maximize the log likelihood. \par
\begin{equation}
\begin{aligned}
LL(W) &= log\prod_{i=1}^m \sigma(z)^{y_i} (1 - \sigma(z)^{1 - y_i} ) \\ &= \Sigma_{i=1}^{m} (y_i \sigma(z) + (1 - y_i)(1 - \sigma(z)))
\end{aligned}
\end{equation} \par
When using mini batch gradient descent (MBGD) algorithm, we want to minimize something, so we minimize the negative log likelihood. \par

\begin{equation}
\label{eqa: NLL}
NLL(W) = -\Sigma_{i=1}^{m} (y_i \sigma(z) + (1 - y_i)(1 - \sigma(z)))
\end{equation} \par
Equation~\ref{eqa: NLL} is also called binary cross entropy.

\subsection{SVM and Hinge Loss}
Given dataset $ D=\{x_{i}, y_{i}\}_{i=1}^{m}$, where $x_{i} \in \mathbf{R}^{n}$ and $y_{i} \in \{-1, 1\}$. SVM is based on linear regression. When given a data point, it wants: \par

\begin{equation}
\label{eqa:svm_two}
\left\{
\begin{array}{lr}
w^{T}x + b \geq 1,~&if~y = 1  \\
w^{T}x + b \leq -1,~&if~y = -1 \\
\end{array}
\right.
\end{equation} \par
Equation~\ref{eqa:svm_two} can also be written to:
\begin{equation}
	y(w^{T}x + b) \geq 1
\end{equation} \par
Other than make the classification correct, SVM also wants to make the classification boundary as safe as possible. To achieve that, it wants to make the margin between the positive class and negative class as large as possible. \par
The positive boundary of SVM is $w^Tx + b = 1$ and negative boundary of SVM is $w^Tx + b = -1$. So the margin between this two boundary is: \par
\begin{equation}
	\label{eqa:margin}
	Margin = \frac{2}{||w||}
\end{equation} \par
In practice, instead of maximize Equation~\ref{eqa:margin}, we want to minimize: \par
\begin{equation}
	\frac{w^Tw}{2}
\end{equation} \par
So, the optimization problem for SVM is:
\begin{equation}
	\begin{aligned}
	\label{eqa:svm_optimization}
	min_{w, b}~&\frac{w^Tw}{2} \\
	s.t.~&y_i(w^Tx_i + b) \geq 1, i=1, 2, ..., m
	\end{aligned}	
\end{equation} \par

Moreover, data point may not be linearly separable. Even the data point is linear separable, the outliers in dataset will hurt the performance of SVM. So, SVM introduce slack variable $\xi_i $ for each data point, which represents how much data point $i$ is on the wrong side of margin boundary. \par
1. if $\xi_i = 0$, it is ok. \par
2. if $ 0 < \xi_i < 1 $, the data point is correctly classified, but with a smaller margin than $\frac{1}{||w||}$. \par
3. if $\xi_i > 1$, the data point is incorrectly classified. \par
Then the optimization problem becomes: 
\begin{equation}
	\begin{aligned}
	\label{eqa:svm_soft_optimization}
	min_{w, b}~&\frac{w^Tw}{2} + \frac{C}{m}\Sigma_{i=1}^m \xi_i \\
	s.t.~&y_i(w^Tx_i + b) \geq 1 - \xi_i, i=1, 2, ..., m
	\end{aligned}	
\end{equation} \par
As $ \xi_i \geq 1 - y_i(w^Tx_i + b), i=1, 2, ..., m$, so
\begin{equation}
\label{eqa:xi}
\xi_i = 
\left\{
\begin{array}{lr}
	1 - y_i(w^{T}x_i + b),&if~1 - y_i(w^{T}x_i + b) \geq 0  \\
	0,&if~1 - y_i(w^{T}x_i + b) < 0  \\
\end{array}
\right.
\end{equation} \par
So \par
\begin{equation}
	\label{eqa:hinge_loss}
	\xi_i = max(0, 1 - y_i(w^Tx_i + b))
\end{equation} \par
Equation~\ref{eqa:hinge_loss} is the Hinge Loss (HL). Finally the optimization problem for SVM becomes
\begin{equation}
\label{eqa:svm_final_optimization}
min_{w, b}~\frac{w^Tw}{2} + \frac{C}{m}\Sigma_{i=1}^m \max(0, 1 - y_i(w^Tx_i + b))
\end{equation} \par
We do a small modification for Equation~\ref{eqa:svm_final_optimization}, so that we can see the role of $C$ in SVM.
\begin{equation}
	\begin{aligned}
	min_{w, b}~\lambda w^Tw + \frac{1}{m}\Sigma_{i=1}^m \max(0, 1 - y_i(w^Tx_i + b)) \\
	where~\lambda = \frac{1}{2C}
	\end{aligned}
\end{equation} \par
Here, we can see that $\lambda$ e.g. $\frac{1}{2C}$ is the regularizer, like Ridge Regression (RR). So, the magnitude of $W$ is in direct proportion to $C$.

\section{Experiments}
\subsection{Dataset}
We conduct all the experiments on a9a dataset, which has total 48842 examples and each example's form is $(x, y)$, where $x \in \mathbf{R}^{123}$ and $y \in \{0, 1\}~(in~LR)~and~\{-1, +1\}~(in~SVM)$. We split the dataset into three part: 26049 training examples, 6512 validation examples and 16281 test examples. We normalized the $X_{train}, X_{val}, X_{test} $ by using the mean and variance of $X_{train}$.

\subsection{Implementation}
We implement the LR, SVM, optimizers, initializers and loss functions using python and mainly rely on the numpy package.

\subsection{Hyper-parameter C in SVM}
In this section, we explore the influence of C in SVM. We set $log_{10}^{C}$ in a coarse-grained range e.g. -2, -1, 0, 1, 2, 3, 4, and compare $loss_{val}$,  $accuracy_{val}$ and $l_2norm$ of $w$, where $loss =  \frac{w^Tw}{2} + \frac{C}{m}\Sigma_{i=1}^m \max(0, 1 - y_i(w^Tx_i + b))$ and $ l_2norm(w) = \sqrt{\Sigma_{i=1}^n w_i}$. We zero initilize $w$ and $b$, use SGD as optimizer with learning rate 0.001, set epoch to 200 and batch size to 64. \par
From Fig~\ref{fig:C_loss}, we can see that, when $C$ is very small, e.g. smaller than 1, the gradient descent process is very unstable and the loss is higher than other. As $C$ goes up, the loss decreases, e.g. from 1 to 10. But when $C$ gets larger than 10, the loss become slightly bigger, but still better than small $C$, like 0.01 and 0.1. So, we can draw the conclusion that, bigger $C$ usually better than small $C$, and the best C is in the middle, which is not very big but also not very small. As we can see from Fig~\ref{fig:C_loss}, $C = 10$ gets the lowest loss.   \par
From Fig~\ref{fig:C_acc}, we can see that, when $C$ is very small, e.g. 0.01, the accracy in both training set and validation set is much lower than others. Small $C$ like 0.01 and 0.1 get worse performance comparing with others. In training set, the accracy is very similar when $C$ bigger than 1. In validation set, $C = 10$ gets slightly better performance comparing with others. Best validation performance under each $C$ setting can be found in Table~\ref{tab:C_performance}. \par
From Fig~\ref{fig:c_mag}, we can see that, as $C$ goes up, $l_2norm(w)$ goes up except $C = 100$. As said above, $\frac{1}{2C}$ can be viewed as regurilizer, so when $C$ is bigger, the smaller of its impact on regurization, so the weight maginitude goes up. \par

>>>>

\usepackage{mathbbol}
\makeatletter
\newif\if@restonecol
\makeatother
\let\algorithm\relax
\let\endalgorithm\relax
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}%[ruled,vlined]{
\usepackage{algpseudocode}
\usepackage{amsmath}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}  % Use Input in the format of Algorithm
\renewcommand{\algorithmicensure}{\textbf{Output:}} % Use Output in the format of Algorithm 
% Other popular packages for formatting tables and equations include:

@@

% Define document title and author
	\title{Adaboost for face image classification}
	\maketitle

% Write abstract here
\begin{abstract}
In this report, we solve the face/non-face classification problem by adaboost with Decision Tree Classifier (DTC) as weak learner. \\
We perform experiments on three aspects: \\
1. Weak learner's performance effect. \\
2. Number of weak learner's exploration. \\
\end{abstract}

% Each section begins with a \section{title} command
\section{Introduction}
	% \PARstart{}{} creates a tall first letter for this first paragraph
\PARstart{A}{daboost} short for "Adaptive Boosting",  is an ensemble method that combine a bunch of weak learners into a strong learner which outperforms all the weak learners. It's wildly used in academic and industry community, which make it very important for us to fully explore it. We will explore the the effect of weak learner's performance and the performance with different number of weak learner.

% Main Part
\section{Methods and Theory}
In this part, we define Adaboost algorithm.
\subsection{Adaboost}
Given dataset $ D=\{x_{i}, y_{i}\}_{i=1}^{m}$, where $x_{i} \in \mathbf{R}^{n}$. $y_{i} \in \{-1, 1\}$ and weak Learner $L(x) \in \{-1, 1\} $, Adaboost algorithm can be found in Algorithm~\ref{alg:adaboost}. \par

\begin{algorithm}
	\label{alg:adaboost}
	\caption{Adaboost}
	\KwIn{Training Dataset $ D=\{x_{i}, y_{i}\}_{i=1}^{m}$, num weak learner $N$, weak learner $L$}
	\KwOut{Final learner $H(x)$}
	$w_1(i) \gets \frac{1}{m}$ where i = 1, 2, ..., m \\
	\For{$idx = 1, 2, ..., N$}
	{
		$h_{idx}(x) = L(D, w_{idx})$ \\
		$\epsilon_{idx} = \Sigma_{i=1}^m w_{idx}(i) \mathbb{1}(h_{idx}(x_i) \neq y_i) $ \\
		$\alpha_{idx} = \frac{1}{2} \log{\frac{1 - \epsilon_{idx}}{\epsilon_{idx}}}$ \\
		$w_{idx+1}(i) = w_{idx}(i) * exp(-1 * \alpha_{idx} * y_i * h_{idx}(x_i))$ where i = 1, 2, .., m \\
		$w_{idx+1} = \frac{w_{idx+1}}{\Sigma_{i=1}^m w_{idx}(i)}  $ \\
	}
	return $H(x) = \Sigma_{idx=1}^N \alpha_{idx} h_{idx}(x)$
\end{algorithm} \par

We can see that, in each iteration, the examples that correctly classified by previous weak learner, will get a smaller weight, so that current learner can pay more attention to the misclassified examples.

\section{Experiments}
\subsection{Dataset}
We conduct all the experiments on self constructed face/non-face dataset, which has 500 256x256 face images and 500 32x32 non-face images. All images have 3 channels. We preprocess all the images into 24x24 gray scale images. We then convert them by using NPD features. Finally, each example's form is $(x, y)$, where $x \in \mathbf{R}^{165600}$ and $y \in \{-1, +1\}$, where $+1$ means face and $-1$ means non-face. We randomly split the dataset into three part: 700 training examples, 100 validation examples and 200 test examples. $X_{train},~X_{val}~and~X_{test} $ are divided by 256 for scale purporse.

\subsection{Implementation}
We implement the Adaboost using python and mainly rely on the numpy package.

\subsection{Weak learner's performance's effect.}
In this section, we explore the influence of weak learner's performance in Adaboost. We set number of weak learner to five weak learner due to computation limitation. \par 
In this section, we explore the influence of C in SVM. We set $log_{10}^{C}$ in a coarse-grained range e.g. -2, -1, 0, 1, 2, 3, 4, and compare $loss_{val}$,  $accuracy_{val}$ and $l_2norm$ of $w$, where $loss =  \frac{w^Tw}{2} + \frac{C}{m}\Sigma_{i=1}^m \max(0, 1 - y_i(w^Tx_i + b))$ and $ l_2norm(w) = \sqrt{\Sigma_{i=1}^n w_i}$. We zero initilize $w$ and $b$, use SGD as optimizer with learning rate 0.001, set epoch to 200 and batch size to 64. \par
From Fig~\ref{fig:C_loss}, we can see that, when $C$ is very small, e.g. smaller than 1, the gradient descent process is very unstable and the loss is higher than other. As $C$ goes up, the loss decreases, e.g. from 1 to 10. But when $C$ gets larger than 10, the loss become slightly bigger, but still better than small $C$, like 0.01 and 0.1. So, we can draw the conclusion that, bigger $C$ usually better than small $C$, and the best C is in the middle, which is not very big but also not very small. As we can see from Fig~\ref{fig:C_loss}, $C = 10$ gets the lowest loss.   \par
From Fig~\ref{fig:C_acc}, we can see that, when $C$ is very small, e.g. 0.01, the accracy in both training set and validation set is much lower than others. Small $C$ like 0.01 and 0.1 get worse performance comparing with others. In training set, the accracy is very similar when $C$ bigger than 1. In validation set, $C = 10$ gets slightly better performance comparing with others. Best validation performance under each $C$ setting can be found in Table~\ref{tab:C_performance}. \par
From Fig~\ref{fig:c_mag}, we can see that, as $C$ goes up, $l_2norm(w)$ goes up except $C = 100$. As said above, $\frac{1}{2C}$ can be viewed as regurilizer, so when $C$ is bigger, the smaller of its impact on regurization, so the weight maginitude goes up. \par
